Dirac neutrino mass generation from Majorana messenger

### 2019-08-02 draft.tex
beta version of introduction

\section{Introduction}
\label{sec:intro}
%Dirac or Majorana

The interpretation of neutrino experimental data in terms of neutrino
oscillations is compatible with both Majorana or Dirac neutrino masses
\cite{Tanabashi:2018oca}. The former possibility has received the most
attention but, given the lack of signals in neutrinoless double beta
decay experiments
\cite{KamLAND-Zen:2016pfg,Agostini:2018tnm,Aalseth:2017btx,Alduino:2017ehq,Albert:2017owj,Arnold:2016bed},
the latter cannot be dismissed.
%Dirac case
If neutrinos are Dirac particles, the
Standard Model (SM) particle content must be extended with
right-handed neutrinos,
%with cosmology implications
which can increase the effective number of
light neutrinos, $N_{\text{eff}}$, until 6. Therefore, to be compatible with the
current cosmological restrictions on $N_{\text{eff}}$, the interactions
of the extra right-handed neutrinos with the primordial plasma must be
highly suppressed.

%Smallness of neutrinos for both Dirac and Majorana
On the other hand, To give small masses to at least two Majorana or
Dirac neutrinos, as required to explain the neutrino oscillation
experiments~\cite{Ahmad:2002jz, Fukuda:1998mi},
%focus in seesaw type I
the seesaw mechanism
with heavy fermions is usually invoked.
%tree-level case
For the tree-level type-I
seesaw we can have either light Majorana neutrinos with heavy Majorana
mediators~\cite{Minkowski:1977sc,Yanagida:1979as,GellMann:1980vs,Mohapatra:1979ia}
or light Dirac neutrinos with heavy Dirac
mediators~\cite{Roncadelli:1983ty,Roy:1983be,Gu:2007mc,Ma:2014qra}.
%one-loop case
The
radiative type-I seesaw includes both~\cite{} possibilities~\cite{},
%new possibility in radiative case
but now it is also possible to have light Majorana neutrinos with

heavy Dirac mediators~\cite{}.
%New in this work
In this work we want to explore the
possibility to build a simple Dirac radiative type-I seesaw  model with 
heavy Majorana mediators.
%Old idea but first realized here
It is worth noticing that this idea have
been already illustrated in an extension of the minimal supersymmetric
standard model~\cite{Demir:2007dt} but without show any explicit
solution.

%Features of the new propossed solution
In general, solutions for light Dirac neutrino masses require a continous symmetry to guarantee their Diracness. This symmetry is usally identified as the $\operatorname{U}(1)_{B-L}$. In models for light Dirac neutrinos, with some explanation for the smallness of
their masses, usually  adhoc discrete symmetries are invoked to forbbid
tree level Dirac or Majorana mass terms for the light right-handed neutrinos~\cite{Han:2018zcn,Wang:2017mcy}.
%tree-level without adhoc symmetries
The tree-level Dirac type-I seesaw with
proper choices for the $\operatorname{U}(1)_{B-L}$ charges have been shown to be consistent without require extra ad-hoc discrette symmetries~\cite{}.
%one-loop 
In recent works, it has been shown that even at one-loop it is possible to have  $\operatorname{U}(1)_{B-L}$ as the only extra symmetry beyond the SM~\cite{Calle:2018ovc, Bonilla:2018ynb, Saad:2019bqf}\footnote{
%extra-loops
  For extensions with only extra scalars, solutions have been found with two and three loops}.
%conexion with dark matter
In this radiative cases, as a bonus,  the new scalars and fermions circulating the loop, can be dark matter candidates with the stability of the ligthest particle guarantee by the  very same continous symmetry.
% remark
We focus here in solutions for the radiative Dirac type-I seesaw, with
only an extra symmetry responsible for the Diracness of the light neutrinos, the absence of any tree-level contribution and the existence of a dark sector constituyed by the particles circulating the loop.

%conexion with DM
In fact, another evidence that the SM is not a complete theory is the missing mass in the universe, which is known as dark matter (DM). The main proposals that explain DM as a particle are given in Ref.~\cite{Bertone:2004pz}. We will make use of the scotogenic model~\cite{Ma:2006km}. In scotogenic model the neutrino masses and the DM have common explanation. The neutrino masses are generated by one-loop effects, where the candidates for DM are mediators in the loop.

In this work we propose a Dirac radiative type-I seesaw,
with Majorana fermion of the mediators in the loop.. This mediators can to be DM candidate. The rest of the paper is organized as follows. In the next section we present the model and the particle content. In the section~\ref{sec:ScaMassSpect} we study the scalar mass spectrum after spontaneous symmetry breaking. In section~\ref{sec:Neutrinos} We present a mechanism to generate Dirac neutrino masses. In section~\ref{sec:LFV} we show the lepton flavor violation (LFV)experimental  constraints. Finally, in the section~\ref{sec:DM} we present the viable DM candidates.

>>>>

\section{Introduction}
\label{sec:intro}
%Dirac or Majorana
%TODO: check ad-hoc symmetries
The interpretation of neutrino experimental data in terms of neutrino
oscillations is compatible with both Majorana or Dirac neutrino masses
\cite{Tanabashi:2018oca}. The former possibility has received the most
attention but, given the lack of signals in neutrinoless double beta
decay experiments
\cite{KamLAND-Zen:2016pfg,Agostini:2018tnm,Aalseth:2017btx,Alduino:2017ehq,Albert:2017owj,Arnold:2016bed},
the latter cannot be dismissed.
% Dirac case
If neutrinos are Dirac particles, the Standard Model (SM) particle
content must be extended with right-handed neutrinos,
% with cosmology implications
which can increase the effective number of light neutrinos,
$N_{\text{eff}}$, until 6. Therefore, to be compatible with the
current cosmological restrictions on $N_{\text{eff}}$, the
interactions of the extra right-handed neutrinos with the primordial
plasma must be highly suppressed.

%Smallness of neutrinos for both Dirac and Majorana
On the other hand, to give small masses to at least two Majorana or
Dirac neutrinos, as required to explain the neutrino oscillation
experiments~\cite{Ahmad:2002jz, Fukuda:1998mi},
%focus in seesaw type I
the seesaw mechanism with heavy fermions is usually invoked.
%tree-level case
For the tree-level type-I seesaw we can have either light Majorana
neutrinos with heavy Majorana
mediators~\cite{Minkowski:1977sc,Yanagida:1979as,GellMann:1980vs,Mohapatra:1979ia}
or light Dirac neutrinos with heavy Dirac
mediators~\cite{Roncadelli:1983ty,Roy:1983be,Gu:2007mc,Ma:2014qra}.
%one-loop case
The radiative type-I seesaw includes both~\cite{Ma:2006km}
possibilities~\cite{Farzan:2012sa},
%new possibility in radiative case
but now it is also possible to have light Majorana neutrinos with
heavy Dirac mediators~\cite{Ma:2013yga}.
%New in this work
In this work we want to explore the possibility to build a simple
Dirac radiative type-I seesaw model with heavy Majorana mediators.
%Old idea but first realized here
It is worth noticing that this idea have been already illustrated in
an extension of the minimal supersymmetric standard
model~\cite{Demir:2007dt} but without show any explicit solution.

%Features of the new propossed solution
In general, solutions for light Dirac neutrino masses require a
continuous symmetry to guarantee their Diracness. This symmetry is
usually identified as the $\operatorname{U}(1)_{B-L}$.  Additionally,
ad hoc discrete symmetries are invoked to forbid tree level Dirac or
Majorana mass terms for the light right-handed
neutrinos~\cite{Roncadelli:1983ty,Han:2018zcn,Wang:2017mcy}. 
%tree-level without adhoc symmetries
However, tree-level Dirac type-I seesaw with proper choices for the
$\operatorname{U}(1)_{B-L}$ charges have been shown to be consistent
without require any extra ad-hoc discrete symmetry~\cite{Ma:2014qra}.
%one-loop 
In recent works, it has been shown that even for one-loop Dirac
neutrino masses, it is possible to have $\operatorname{U}(1)_{B-L}$ as
the only extra symmetry beyond the SM~\cite{Calle:2018ovc,
  Bonilla:2018ynb, Saad:2019bqf}\footnote{
%extra-loops
  For extensions with only extra scalars, solutions have been found with two and three loops~\cite{Saad:2019bqf}}.
%conexion with dark matter
As a bonus in this case, the new scalars and fermions circulating the
loop can be dark matter candidates with the stability of the lightest
of them guaranteed by the very same continuous symmetry.
% remark
We focus here in solutions for the radiative Dirac type-I seesaw with
Majorana mediators, which have only an extra symmetry responsible for the
Diracness of the light neutrinos, the absence of any tree-level
contribution and the existence of a dark sector constituted by the
particles circulating the loop.

%conexion with DM
In fact, another evidence that the SM is not a complete theory is the
missing matter content of the universe, which is known as dark matter
(DM).
% DM as particle
The main proposals that explain DM as a particle are given in
Ref.~\cite{Bertone:2004pz}.
% too many possibilities
Without evidence of DM as a particle there is not a clear path to pin
out the DM properties nor the possible heavier companions of some
extended sector. 
% connect with other phenomena
Linking the dark sector to other specific phenomenology allows to
reduce the arbitrariness in the model building.
% loop
In our construction, the dark sector is related to the heavy sector
responsible of the lightness of the neutrinos and the same symmetry
which guarantees the lightness of the Dirac neutrinos is the
responsible of the stability of the lightest dark particle (LDP).
% new particle prediction
The number of specific models is restricted.

% We will make use of the scotogenic model~\cite{Ma:2006km}. In
% scotogenic model the neutrino masses and the DM have common
% explanation. The neutrino masses are generated by one-loop effects,
% where the candidates for DM are mediators in the loop.
%Conclusion:
In this work we propose a Dirac radiative type-I seesaw, with Majorana
fermion of the mediators in the loop. This mediators can include a DM
candidate.

The rest of the paper is organized as follows. In the next section we present the model and the particle content. In the section~\ref{sec:ScaMassSpect} we study the scalar mass spectrum after spontaneous symmetry breaking. In section~\ref{sec:Neutrinos} We present a mechanism to generate Dirac neutrino masses. In section~\ref{sec:LFV} we show the lepton flavor violation (LFV)experimental  constraints. Finally, in the section~\ref{sec:DM} we present the viable DM candidates.
